---
title: "abmR: An R Package for Agent-based Model Analysis of Large-scale Movements Across Taxa"
author: "Benjamin Gochanour, Javi Fernandez Lopez, Andrea Contina"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# I. Getting Started

## Installation

To use `abmr`, you must first install it from Github using `devtools` 
and load the library:
```{r}
#devtools: install_github("bgoch5/move-model",subdir="abmr")
library(abmr,quietly=TRUE,warn.conflicts=FALSE)
```
While this package is still in development, it will be updated frequently, so please
be sure to update it frequently.

## Requirements for Environmental Data Rasters and Species objects

The modeling functions discussed below require several input objects to work, which
we will discuss here.

# II. Running Agent-based models

## `moveSIM`

## `energySIM`

# III. Visualizing Results

The output from 'moveSIM` and `energySIM` are simple dataframes, so one may easily
visualize their results using custom codes. However, we have provided the following functions that include a set of default plots that you may find useful.

## `moveVIZ`

## `energyVIZ`

# IV. Getting Help

Need help? Have suggestions to make `abmr` better? If so, please open an issue or
pull request on Github (https://github.com/bgoch5/move-model/tree/master/abmr), or drop me an email at ben.gochanour\@ou.edu.
